---
title: "Homework 1"
author: "Your Name Here"
date: "Due Oct 9, 2023"
format: pdf
---

## Question 1: Expectations

#### 1.a) Define both and explain the difference between (1) the expectation of a random variable and (2) the sample average.

...add answer here...


## Question 2: LLN & CLT

#### 2.a) Plot the density of a Beta(5,2) distribution over it's domain [0,1]. Make a publication-quality plot by changing any unwanted default plotting behavior and by adding relevant titles and labels. The first parameter of the Beta distribution is often labeled $\alpha$ and the second $\beta$. The `dbeta()` and related functions in R label these parameters as `shape1` and `shape2`.

```{r}
# ...add answer here...
```

#### 2.b) State the Law of Large Numbers as simply as you can.

...add answer here...

#### 2.c) Set the seed to the value 1234 (`set.seed(1234)`). Then take 1,000 random draws from the Beta(5,2) distribution using `rbeta()`. Calculate a running sample average. Specifically: calculate $\bar{X}_1 = x_1$, then calculate $\bar{X}_2 = (1/2) \sum_{i=1}^2 X_i$, then calculate $\bar{X}_3 = (1/3) \sum_{i=1}^3 X_i$. Continue until you have calculated $\bar{X}_{1000} = (1/1000) \sum_{i=1}^{1000} X_i$. Create a scatterplot with the values 1--1,000 on the horizontal axis and the 1,000 cumulative average values of $\bar{X}_i$ for $i=1,\ldots,1000$ you calculated on the vertical axis. Compare your value for $\bar{X}_{1000}$ to the $\mathbb{E}[X]=\alpha/(\alpha+\beta)=5/7 = 0.7143$.

```{r}
# ...add answer here...
```


#### 2.d) State the Central Limit Theorem as simply as you can.

...add answer here...

#### 2.e) Make two plots. For the first plot, take D=10 draws from the Beta(5,2) distribution and calculate the sample average. Repeat the process of taking D=10 draws and finding the sample average R=10,000 times. Plot a histogram of the 10,000 sample averages. For the second plot, repeat the process with D=100 draws. These two histograms are called "sampling distributions." 

```{r}
# ...add answer here...
```


## Question 3: Linear Algebra

**Suppose $\mathbf{X}$ and $\mathbf{Y}$ are defined as follows.**

$$ \mathbf{X} = \begin{bmatrix} 1 & 1 \\ 1 & 4 \\ 1 & 5 \\ 1 & 8 \end{bmatrix}  \hspace{5em} \mathbf{y} = \begin{bmatrix} 6 \\ 5 \\ 3 \\ 2 \end{bmatrix} $$

#### 3.a) What is the rank of $\mathbf{X}$? Provide a brief (approx 1 sentence) explanation. Check your work via R code using `Matrix::rankMatrix()`.

...add answer here...

```{r}
# ...check answer here...
```

#### 3.b) Calculate $\mathbf{X}'\mathbf{X}$. Use the `bmatrix` environment in Latex to typeset your answer. Check your work via R code. 

...add answer here...

```{r}
# ...check answer here...
```

#### 3.c) What is the rank of $\mathbf{X}'\mathbf{X}$? Provide a brief explanation. Check your work via R code.

...add answer here...

```{r}
# ...check answer here...
```

#### 3.d) Find $(\mathbf{X}'\mathbf{X})^{-1}$ "by hand" (as you would with paper and pencil) using the approach outlined on slide 47 of the Class 1 slides. Check your work via R code.

...add answer here...

```{r}
# ...check answer here...
```

#### 3.e) What is the rank of $(\mathbf{X}'\mathbf{X})^{-1}$? Provide a brief explanation. Check your work via R code.

...add answer here...

```{r}
# ...check answer here...
```

#### 3.f) Calculate $\mathbf{X}'\mathbf{y}$. Check your work via R code.

...add answer here...

```{r}
# ...check answer here...
```

#### 3.g) Use your results from 3d and 3f to calculate "by hand" $(\mathbf{X}'\mathbf{X})^{-1}\mathbf{X}'\mathbf{y}$.  Check your work via R code.

...add answer here...

```{r}
# ...check answer here...
```




